---
title: "Assignment 7"
author: "Michelle"
date: "4/28/2020"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#First get packages we need
```{r, eval=FALSE}
install.packages("yearn")
install.packages("OUwie")
install.packages("nloptr")
install.packages("lattice")
install.packages("phytools")
```

```{r}
library(ape)
library(geiger)
library(OUwie)
library(nloptr)
library(lattice)
library (corHMM)
library(phangorn)
library(openxlsx)
library(rotl)
library(phytools)
```


```{r}
hollies <- get_study_tree("ot_1984", "tree1")
hollies.study.metadat<-get_study_meta("ot_1984")
hollies.tree.ids<-get_tree_ids(hollies.study.metadat)
plot(hollies, type="fan", cex = 0.4)
```

#Now get the tree and data. For these exercises, knowing uncertainty in your measurements can also be important. #(remember for homework to change eval=FALSE to eval=TRUE).
```{r}
hollies <- get_study_tree("ot_1984", "tree1")
print (hollies)
hollies.study.metadat<-get_study_meta("ot_1984")
hollies.tree.ids<-get_tree_ids(hollies.study.metadat)
plot(hollies, cex = 0.4)

setwd("~/phylometh_exercises-/data\ set\ 6\ and\ 7")

hollies.data <- read.xlsx (xlsxFile = "hollies.xlsx", startRow = 2)
head(hollies.data )

#Take the first 2 names and anytime there is an "_" it will split and put it together.
GetGenusSpeciesOneTaxon <- function(x) {
  x <- gsub("_", " ", x)
  x <- strsplit(x, "\\s+")[[1]]
  return(paste0(x[1:2], collapse="_"))
}
GetGenusSpeciesManyTaxa <- function(x) {
  return(unname(sapply(x, GetGenusSpeciesOneTaxon)))
}
# To get ride of the space and replace it with an "_"
hollies.data$species <- GetGenusSpeciesManyTaxa(gsub(" ", "_", hollies.data$species))

head(hollies.data )

#To read csv files
#discrete.data <- read.csv(file= "hollies.xlsx", stringsAsFactors=FALSE)
```


```{r}
#To get the fruit.diameter.(mm) data
#hollies.traits <- hollies.data[,c("species", "fruit.diameter", "colour")]
#print(head(hollies.traits))
print(colnames(hollies.data))
hollies.fruit.diameter <- hollies.data$fruit.diameter
names(hollies.fruit.diameter) <- hollies.data$species
print(head(hollies.fruit.diameter))
print(hollies.fruit.diameter)
hollies.colour <- hollies.data$colour
names(hollies.colour) <- hollies.data$species
print(head(hollies.colour))
print(hollies.colour)
hollies.colour <- hollies.colour[!grepl("na", hollies.colour)]
print(hollies.colour)
```

```{r}
#Omit NA from data set
hollies.fruit.diameter <- hollies.fruit.diameter[!grepl("na", hollies.fruit.diameter)]
print(hollies.fruit.diameter)
```

#A function to clean data, make sure taxon names match between tree and data, etc.
#fruit diameter
```{r, eval=TRUE}
CleanData_cont <- function(phy, data) {
  data <- log(data) #because we can't have negative fruit size.
 #result=treedata(phy, data,sort = TRUE, warnings = FALSE)
  #return(result) You can used the two above functions (results or return) or the function below
  return(treedata(phy, data,sort = TRUE, warnings = FALSE))
}
h_cleanedF<-CleanData_cont(hollies, hollies.fruit.diameter)
print(h_cleanedF)
```

#fruit color
```{r, eval=TRUE}
CleanData_discrete <- function(phy, data) {
 #result=treedata(phy, data,sort = TRUE, warnings = FALSE)
  #return(result) or the function below
  return(treedata(phy, data,sort = TRUE, warnings = FALSE))
}
# three objects: original tree, colour, diameter
# first cleaning: tree_diameter_cleaned, diameter_tree_cleaned, original colour
# second cleaning: tree_diameter_and_colour_cleaned, diameter_tree_cleaned, colour_diameter_and_tree_cleaned
# third cleaning: tree_diameter_and_colour_cleaned, diameter_tree_and_colour_cleaned, colour_diameter_and_tree_cleaned

#Tree: A, B, C, D, E, no F
# Diameter: A, C, D, E, F, no B
# Colour: B, C, D, E, F, no A


# Tree + Colour + Diameter, all with C, D, E only


# Tree diameter cleaning:
# Tree A, C, D, E  CLEAN 
# Diameter A, C, D, E CLEAN
# Colour, B, C, D, E, F

# Colour cleaning
# Tree C, D E CLEAN
# Diameter A, C, D, E
# Colour C, D. E CLEAN

# clean both 
# TREE C, D, E CLean
# Diameer C D E
# Colour C D E CLEAN
h_cleanedC<-CleanData_discrete(h_cleanedF$phy, hollies.colour)
print(str(h_cleanedC))
```

#Third cleaned
```{r}
h_cleanedBoth <- treedata(h_cleanedC$phy, h_cleanedF$data )

clean.tree <- h_cleanedBoth$phy
clean.colour <- h_cleanedC$data
clean.diameter <- h_cleanedBoth$data
```


#A function to plot data. Look at phytools::contMap(). This is all part of checking: do your data all seem #sensible? #LOOK AT IT.
```{r, eval=TRUE}
VisualizeData_Discrete <- function(phy=phy, data=data) {
  dotTree(phy,data,length=10,fsize=0.5,lwd=2)
}
VisualizeData_Discrete(phy=clean.tree, data=clean.diameter)
```

```{r, eval= TRUE}
print(paste("This tree has a polytomy? ", is.binary(h_cleanedF$phy)))
print(h_cleanedF$data[,1])
VisualizeData_Conti <- function(phy, data) {
 pretty_pic <- phytools::contMap(tree = phy, x = data, fsize = c(.2,1), lwd = 2)
 #polytomy <- is.binary(phy = phy)
 pretty_pic
}
VisualizeData_Conti(phy=clean.tree,data=clean.diameter[,1])
```
 
#First, start basic. What is the rate of evolution of your trait on the tree?
```{r, eval=TRUE}
BM1 <- geiger::fitContinuous(phy=clean.tree, dat=clean.diameter, model="BM")
print(str(BM1))
print(paste("The rate of evolution is", BM1$opt$sigsq, "in units of", "(log mm)^2 / millions of years"))
```

#Important: What are the rates of evolution? In what units?
```{r,eval=TRUE}
OU1 <- fitContinuous(phy=clean.tree, dat=clean.diameter, model="OU")
par(mfcol=c(1,2))
plot(h_cleanedF$phy, show.tip.label=FALSE)
axisPhylo()
print(OU1)
print("now doing str")
print(str(OU1))
ou.tree <- rescale(h_cleanedF$phy, model="OU", OU1$opt$alpha)
plot(ou.tree, show.tip.label=FALSE)
axisPhylo() #see how long the tree is
```

#Compare trees
```{r, eval= TRUE}
AIC.BM1 <- BM1$opt$aic
AIC.OU1 <- OU1$opt$aic
delta.AIC.BM1 <-AIC.BM1-min(AIC.BM1,AIC.OU1)
delta.AIC.OU1 <- AIC.OU1-min(AIC.OU1,AIC.BM1)
print(delta.AIC.BM1)
print(delta.AIC.OU1)
```

#First, we need to assign regimes. The way we do this is with ancestral state estimation of a discrete trait. We #can do this using ace() in ape, or similar functions in corHMM or diversitree. Use only one discrete char
```{r, eval= TRUE}
one.discrete.char <- clean.colour
reconstruction.info <- ace(one.discrete.char, clean.tree, type="discrete", method="ML", CI=TRUE)
best.states <- colnames(reconstruction.info$lik.anc)[apply(reconstruction.info$lik.anc, 1, which.max)]
```

#Now add these labels to your tree
```{r, eval= FALSE}
labeled.tree <- best.states
print(clean.diameter)
clean.diameter.df<-data.frame(species=rownames(clean.diameter), diameter=clean.diameter[,1])
nodeBased.OUMV <- OUwie(clean.tree, clean.diameter.df, model="OUMV", simmap.tree=FALSE, diagn=FALSE)
print(nodeBased.OUMV)
```
#What do the numbers mean?

#Now run all OUwie models:
```{r, eval=FALSE}
models <- c("BM1","BMS","OU1","OUM","OUMV","OUMA","OUMVA")
results <- lapply(models, RunSingleOUwieModel, phy=h_cleanedF$phy, data=h_cleanedF$data)

AICc.values<-sapply(results, "[[", "AICc")
names(AICc.values)<-models
AICc.values<-AICc.values-min(AICc.values)


print(AICc.values) #The best model is the one with smallest AICc score

best<-results[[which.min(AICc.values)]] #store for later

print(best) #prints info on best model
```

```{r, eval=FALSE}
alpha.values<-seq(from= _______________ , to= _______________ , length.out=50)
```

#Keep it simple (and slow) and do a for loop:
```{r, eval=FALSE}
likelihood.values <- rep(NA, length(alpha.values))
for (iteration in sequence(length(alpha.values))) {
    likelihood.values[iteration] <- OUwie.fixed(tree, trait, model="OUMV", alpha=rep(alpha.values[iteration],2), sigma.sq=best$solution[2,], theta=best$theta[,1])$loglik
}

plot(x= _______________ , y= _______________, xlab="_______________", ylab="_______________", type="l", bty="n")
points(x=best$solution[1,1], y=best$loglik, pch=16, col="red")
text(x=best$solution[1,1], y=best$loglik, "unconstrained best", pos=4, col="red")
```

#A rule of thumb for confidence for likelihood is all points two log likelihood units worse than the best value. Draw a dotted line on the plot to show this

```{r, eval=FALSE}
abline(h=_______________, lty="dotted") #Two log-likelihood
```

#Now, let’s try looking at both theta parameters at once, keeping the other parameters at their #MLEs
```{r, eval=FALSE}
require("akima")
nreps<-400
theta1.points<-c(best$theta[1,1], rnorm(nreps-1, best$theta[1,1], 5*best$theta[1,2])) #center on optimal value, have extra variance
theta2.points<-c(best$theta[2,1], rnorm(nreps-1, best$theta[2,1], 5*best$theta[2,2])) #center on optimal value, have extra variance
likelihood.values<-rep(NA,nreps)

for (iteration in sequence(nreps)) {
    likelihood.values[iteration] <- OUwie.fixed(tree, trait, model="OUMV", alpha=best$solution[1,], sigma.sq=best$solution[2,], theta=c(theta1.points[iteration], theta2.points[iteration]))$loglik
}

```

#Think of how long that took to do 400 iterations. Now remember how long the search took (longer).
```{r, eval=FALSE}
likelihood.differences<-(-(likelihood.values-max(likelihood.values)))
```

#We are interpolating here: contour wants a nice grid. But by centering our simulations on the MLE values, we made sure to sample most thoroughly there
```{r,eval=FALSE}
interpolated.points<-interp(x=theta1.points, y=theta2.points, z= likelihood.differences, linear=FALSE, extrap=TRUE, xo=seq(min(theta1.points), max(theta1.points), length = 400), yo=seq(min(theta2.points), max(theta2.points), length = 400))

contour(interpolated.points, xlim=range(c(theta1.points, theta2.points)),ylim=range(c(theta1.points, theta2.points)), xlab="Theta 1", ylab="Theta 2", levels=c(2,5,10),add=FALSE,lwd=1, bty="n", asp=1)

points(x=best$theta[1,1], y=best$theta[2,1], col="red", pch=16)

points(x=trait$X[which(trait$Reg==1)],y=rep(min(c(theta1.points, theta2.points)), length(which(trait$Reg==1))), pch=18, col=rgb(0,0,0,.3)) #the tip values in regime 1, plotted along x axis
points(y=trait$X[which(trait$Reg==2)],x=rep(min(c(theta1.points, theta2.points)), length(which(trait$Reg==2))), pch=18, col=rgb(0,0,0,.3)) #the tip values in regime 2, plotted along y axis
```


```{r,eval=FALSE}
library(phytools)
trait.ordered<-data.frame(trait[,2], trait[,2],row.names=trait[,1])
trait.ordered<- trait.ordered[tree$tip.label,]
z<-trait.ordered[,1]
names(z)<-rownames(trait.ordered)
tree.mapped<-make.simmap(tree,z,model="ER",nsim=1)
leg<-c("black","red")
names(leg)<-c(1,2)
plotSimmap(tree.mapped,leg,pts=FALSE,ftype="off", lwd=1)

simmapBased<-OUwie(tree.mapped,trait,model="OUMV", simmap.tree=TRUE, diagn=FALSE)
print(simmapBased)
print(best)
```

